---
title: "Data Science 2 HW 3"
author: "Amanda Howarth"
date: "4/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = T, message = FALSE, results='hide', warning=FALSE}
library(caret)
library(glmnet)
library(dplyr)
library(MASS) 
library(e1071)
library(pROC) 
library(AppliedPredictiveModeling)
library(ISLR)
library(tidyverse)
library(sqldf)
```


# 1. Produce graphical summaries of the Weekly data
```{r}
data("Weekly")

#Summary of Weekly data
summary(Weekly)

ggplot(Weekly, aes(x = Year, y = Volume, group = Year)) + geom_boxplot() + labs(title = "Total Volume Across Years, 1990 - 2010",
           x = "Year",
           y = "Volume")
```

```{r}
theme1 <- transparentTheme(trans = .4) #creates the themes for the feature plots 
theme1$strip.background$col <- rgb(.0, .6, .2, .2) 
trellis.par.set(theme1)

featurePlot(x = Weekly[, 1:7], 
            y = Weekly$Direction,
            scales = list(x=list(relation="free"), 
                          y=list(relation="free")),
            plot = "density", pch = "|", 
            auto.key = list(columns = 2))
```
Density plots are graphically presented above for seven predictor variables in the dataset (lag1, lag2, lag3, lag4, lag5, volume, year).



```{r}
glm.fit <- glm(Direction~ Lag1 + Lag2 + Lag3 + Lag4 + Lag5 + Volume, 
               data = Weekly,
               family = binomial)

contrasts(Weekly$Direction)
#UP = 1; DOWN = 0

summary(glm.fit)
```

At alpha level of 0.05, only the variable "lag2" appears to be statistically significant in the logistic regression model (p-value = 0.0296). All other predictor variables have p-values greater than 0.05. 
```{r}
glm.probs=predict(glm.fit, type = "response")
glm.probs[1:10]
contrasts(Weekly$Direction)
glm.pred=rep("Down", 1089)
glm.pred[glm.probs>.5]="Up"
table(glm.pred, Weekly$Direction)

(557+54)/1089
# 0.5610652

mean(glm.pred == Weekly$Direction)

```

# Part D
```{r}
roc.glm <- roc(Weekly$Direction, glm.probs)

plot(roc.glm, legacy.axes = TRUE, print.auc = TRUE)
plot(smooth(roc.glm), col = 4, add = TRUE) 
```
The AUC value is 0.554.

# PART E
```{r}
train = (Weekly$Year < 2009)
Weekly.2009=Weekly[!train,]
dim(Weekly.2009)

Direction.2009= Weekly$Direction[!train]
```
There are 104 observations in the test dataset from years 2009 and 2010. 

# LOGISTIC REG
```{r}
glm.fit2 <- glm(Direction~ Lag1 + Lag2, 
               data = Weekly, 
               subset = train, 
               family = binomial)


glm.probs2 =predict(glm.fit2, Weekly.2009, type="response")

glm.pred2 = rep("Down", 104)
glm.pred2[glm.probs2>.5]="Up"
table(glm.pred2, Direction.2009)
mean(glm.pred2 == Direction.2009)
mean(glm.pred2!=Direction.2009)

#using confusion matrix code
confusionMatrix(data = as.factor(glm.pred2),
                reference = Weekly.2009$Direction,
                positive = "Up")

roc.glm2 <- roc(Weekly.2009$Direction, glm.probs2)
plot(roc.glm2, legacy.axes = TRUE, print.auc = TRUE)
plot(smooth(roc.glm2), col = 4, add = TRUE) 
```
The AUC value is 0.556.

#LDA
```{r}
lda.fit = lda(Direction ~ Lag1 + Lag2, data = Weekly, subset = train)

#Creates distributions (histograms) of Z for the two classes
plot(lda.fit)

lda.pred = predict(lda.fit, newdata = Weekly.2009)
head(lda.pred$posterior)
#posterior= predicted class probability

roc.lda <- roc(Weekly.2009$Direction, lda.pred$posterior[,2], 
               levels = c("Down", "Up"))

plot(roc.lda, legacy.axes = TRUE, print.auc = TRUE)
```
The AUC value is 0.557

# QDA
```{r}
qda.fit <- qda(Direction ~ Lag1 + Lag2, 
               data = Weekly,
               subset = train)

qda.pred <- predict(qda.fit, newdata = Weekly.2009)
head(qda.pred$posterior)

roc.lda <- roc(Weekly.2009$Direction, qda.pred$posterior[,2], 
               levels = c("Down", "Up"))

plot(roc.lda, legacy.axes = TRUE, print.auc = TRUE)

```
The AUC value is 0.529. 

```{r}

ctrl <- trainControl(method = "repeatedcv",
                     repeats = 5,
                     summaryFunction = twoClassSummary,
                     classProbs = TRUE)

set.seed(1)
model.knn <- train(Direction ~ Lag1 + Lag2, 
                   data = Weekly, 
                   method = "knn",
                   preProcess = c("center","scale"),
                   tuneGrid = data.frame(k = seq(1,200,by=5)),
                   trControl = ctrl)

ggplot(model.knn)

``` 






